Commutative Character Sheaves

### 2015-08-17 followup.tex
Working on descent to the abelianization
I’m back on the grid!

Let $G$ be a smooth group scheme over a finite field $\Fq$; that is, let $G$ be a group scheme over $\Fq$
for which the structure morphism $G \to \Spec{\Fq}$ is smooth in the sense of [include reference for this definition].
This implies $G \to \Spec{\Fq}$ is locally of finite type, but not that it is of finite type.
We remark that the identity component $G^0$ of $G$ is of finite type over $\Fq$, while the component group scheme
$\pi_0(G)$ of $G$ is an \'etale group scheme over $\Fq$, and both are smooth over $\Fq$.
\todo{Remember to check the matching paragraph in CS for T. It may have an error. The group operations are also smooth, not just the structure map? Hm. It might be okay, actually  . . . \\ -CC}

@@

\section{Passing to the abelianization}

 Let $G$ be a smooth group scheme over $\Fq$ and let $C = [G, G]$ be its commutator subgroup.

@@

 \begin{proposition}
 Every commutative character sheaf $\cs{L}$ on $G$ is equivariant for the left translation action of $C$ on $G$.
 \end{proposition}

 \begin{proof}
Let $n : C \times G \to G$ be the left translation action of $C$ on $G$. 
Then $n = m \circ (j \times \id)$, so
\begin{align*}
n^* \cs{L} 
&= (m \circ (j \times \id))^* \cs{L}\\
&= (j \times \id))^* m^*\cs{L}\\
&\iso (j \times \id))^* (\cs{L} \boxtimes \cs{L})\\
&= (j^*\cs{L}) \boxtimes \cs{L}\\
&= (\cs{L}\vert_C) \boxtimes \cs{L}\\
&= (\EE)_C \boxtimes \cs{L}\\
&= p_2^* \cs{L}.
\end{align*}
 \end{proof}

@@

\begin{proposition}
Every commutative character sheaf on $G$ is equivariant for the conjugation action of $G$ on $G$.
\end{proposition}

@@

\appendix

\section{Descent}

[Since I haven't found a good reference for descent in the non-connected case, I think I should lay out the arguments here. We use descent in this paper to pass from $C$-equivariant local systems on $G$ to local systems on $G\ab$, and thence to characters of $G\ab(\Fq)$. We also used it in our last paper in a different way. It is also lurking behind Weil sheaves ..... So, I'm going to set things up in a very general context. This section will get whittled down to  something more manageable, later.]

>>>>

Let $G$ be a smooth group scheme over a finite field $\Fq$; that is, let $G$ be a group scheme over $\Fq$
for which the structure morphism $G \to \Spec{\Fq}$ is smooth in the sense of [include reference for this definition].
This implies $G \to \Spec{\Fq}$ is locally of finite type, but not that it is of finite type.
We remark that the identity component $G^0$ of $G$ is of finite type over $\Fq$, while the component group scheme
$\pi_0(G)$ of $G$ is an \'etale group scheme over $\Fq$, and both are smooth over $\Fq$.

@@

\section{Descent to the abelianization}

 \begin{theorem}\label{thm:descent}
 Let $G$ be a smooth group scheme over $\Fq$ and let $C = [G, G]$ be its commutator subgroup. 
The quotient $G \to G/C$ exists in the category of smooth group schemes over $\Fq$.
The category of commutative character sheaves on $G$ is equivalent to the category of commutative character sheaves on $G/C$: 
 \[
 \CCS(G) \iso \CCS(G/C).
 \]
 \end{theorem}

\begin{proof}
Set $G\ab \ceq G/C$ and let $q : G \to G\ab$ be the quotient in the category of smooth group schemes over $\Fq$.
Recall the functor $q^* : \CCS(G\ab)\to \CCS(G)$ from  . . . 
The strategy of the proof is indicated by the diagram below. 
\[
\begin{tikzcd}
\CCS(G)& \\
& \CCS_C(G) \arrow{ul}[swap]{\text{forget}} \\
\CCS(G\ab) \arrow{uu}{g^*} \arrow{ur}[swap]{\text{comparison}} & 
\end{tikzcd}
\]
We will first use smooth descent to show that the comparison functor is an equivalence from $\CCS(G\ab)$ to the category $\CCS_C(G)$ of $C$-equivariant commutative character sheaves on $G\ab$, the latter defined below.
Then we will show that the forgetful functor from $\CCS_C(G)$ to $\CCS(G)$ is also an equivalence.

  Every commutative character sheaf $\cs{L}$ on $G$ is equivariant for the left translation action of $C$ on $G$.
Let $n : C \times G \to G$ be the left translation action of $C$ on $G$
and let $p : C \times G \to G$ be projection to the second component.
Then $n = m \circ (j \times \id)$, so
\begin{align*}
n^* \cs{L} 
&= (m \circ (j \times \id))^* \cs{L}\\
&= (j \times \id))^* m^*\cs{L}\\
&\iso (j \times \id))^* (\cs{L} \boxtimes \cs{L})\\
&= (j^*\cs{L}) \boxtimes \cs{L}\\
&= (\cs{L}\vert_C) \boxtimes \cs{L}\\
&= (\EE)_C \boxtimes \cs{L}\\
&= p^* \cs{L}.
\end{align*}

\end{proof}

\begin{lemma}
Every commutative character sheaf on $G$ is equivariant for the conjugation action of $G$ on $G$.
\end{lemma}

@@

@@

@@
